---
title: "BCB570 Assignment 1"
author: "Ashish Jain"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE
)
```

#Question 1:

```{r}

degreeDistribution<-c(2,4,4,4,4,5,4,3,2)
ccDistribution<-c(1,4/6)
hist(degreeDistribution)

```

#Question 2:


#Question 3:


#Question 4:

#Question 5:

#Question 6:
To check for small-world property of the networks, we can do the following:

i). Compute degree distribution of the network.

ii). Try to fit a power law distribution to the degree distribution.

iii). Compare clustering coefficients of the network with that of a random network with
same number of nodes and edges.

iv). Compare average path lengths of the network with that of a random network with 
same number of nodes and edges.

v). For a small world network:


#Question 7:
To check for scale-free property of the networks, we can do the following:

i). Plot degree distribution vs # of nodes (log-log scale) to check for a linear decreasing 
relation.

ii). Plot clustering coefficient vs # of nodes (log-log scale) to check for a constant line.

iii). Check if there are few hubs.

iv). Try to fit a power law distribution to the degree distribution and if p<0.05, the network is not  cale free.

v). Take a subset of the network and check if it retains the properties of the full network.

vi). Expand the network (with more real data or accurately simulated data) and check if it retains the properties of the full network.