---
title: "BCB570 Assignment 3"
author: "Ashish Jain"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE
)
```

#Question 1:

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
library("tidyverse")
library("igraph")

graphDetails<-function(edgeFile)
{
  graph<-as.matrix(read.table((edgeFile)))
  graphGraph1<-graph.data.frame(graph,directed = FALSE)
  graphGraph<-simplify(graphGraph1,remove.multiple = TRUE, remove.loops = TRUE,
  edge.attr.comb = igraph_opt("edge.attr.comb"))
  graphDetails<-c(vcount(graphGraph), ecount(graphGraph), graph.density(graphGraph),diameter(graphGraph), radius(graphGraph), transitivity(graphGraph, type = "average"),transitivity(graphGraph, type = "global"),mean_distance(graphGraph))
  return(list(graphDetails,graphGraph))
}
path<-"https://raw.githubusercontent.com/ashishjain1988/BCB570/master/HW3/"
outputY2H_uniondata<-graphDetails(paste0(path,"Y2H_uniondata.txt"))
outputCCSB_YI1data<-graphDetails(paste0(path,"CCSB_YI1.txt"))
essentialGenes<-read.table(paste0(path,"essentialGenes.txt"))

details<-c("Number of Nodes","Number of Edges","Graph Density","Diameter","Radius","Average Clustering Coefficent","Global Clustering Coefficent","Average Shortest Path")

table<-data.frame(Headings=details,"Y2H-Union"=outputY2H_uniondata[[1]],"CCSB-YI1"=outputCCSB_YI1data[[1]])
table %>% knitr::kable(caption = "PPI Details")

graphUnion<-outputY2H_uniondata[[2]]
graphCCSB<-outputCCSB_YI1data[[2]]

gdetails<-function(graphGraph1)
{
  graphGraph<-simplify(graphGraph1,remove.multiple = TRUE, remove.loops = TRUE,
  edge.attr.comb = igraph_opt("edge.attr.comb"))
  graphDetails<-c(vcount(graphGraph), ecount(graphGraph), graph.density(graphGraph),diameter(graphGraph), radius(graphGraph), transitivity(graphGraph, type = "average"),transitivity(graphGraph, type = "global"),mean_distance(graphGraph))
  return(graphDetails)
}

dgraphUnion <- decompose.graph(graphUnion) # returns a list of three graphs
dgraphCCSB <- decompose.graph(graphCCSB)

lgraphUnion<-dgraphUnion[[1]]
for(i in 1:length(dgraphUnion))
{
  if(vcount(lgraphUnion) < vcount(dgraphUnion[[i]]))
  {
    lgraphUnion <- dgraphUnion[[i]]
  }
}

lgraphCCSB<-dgraphCCSB[[1]]

for(i in 1:length(dgraphCCSB))
{
  if(vcount(lgraphCCSB) < vcount(dgraphCCSB[[i]]))
  {
    lgraphCCSB <- dgraphCCSB[[i]]
  }
}

table<-data.frame(Headings=details,"Y2H-Union"=gdetails(lgraphUnion),"CCSB-YI1"=gdetails(lgraphCCSB))
table %>% knitr::kable(caption = "PPI Details of largest sub-network")

```



```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

##Small world property

getRandomNetworkDetails<-function(graph)
{
  randomNetwork <- erdos.renyi.game(vcount(graph), ecount(graph),type='gnm')
  output1<-average.path.length(graph)
  output2<-average.path.length(randomNetwork)
  output3<-transitivity(graph, type="global")
  output4<-transitivity(randomNetwork, type="global")
  return(c(output1,output2,output3,output4))
}

graphUnionProp<-getRandomNetworkDetails(graphUnion)
graphCCSBProp<-getRandomNetworkDetails(graphCCSB)

details<-c("Average Path Length-Graph","Average Path Length-Random Graph","Clustering Coefficient-Graph","Clustering Coefficient-Random Graph")

table<-data.frame(Features=details,"Y2H-Union"=graphUnionProp,"CCSB-YI1"=graphCCSBProp)
table %>% knitr::kable(caption = "Small-World Property Test")
```

From the table 3, it can be seen that none of the networks satisfies the small-world network properties. None of them have both their clustering coefficient and average path length greater than the random network.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
table<-data.frame(Features=details,"Y2H-Union"=getRandomNetworkDetails(lgraphUnion),"CCSB-YI1"=getRandomNetworkDetails(lgraphCCSB))
table %>% knitr::kable(caption = "Small-World Property Test for largest sub-network")
```

From the table 4, it can be seen that none of the largest sub-network satisfies the small-world network properties. None of them have both their clustering coefficient and average path length greater than the random network.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

##Power law fit
fitUnion <- power.law.fit(degree(graphUnion),9,implementation="plfit") 
fitCCSB <- power.law.fit(degree(graphCCSB),9,implementation="plfit") 

details<-c("Graph Union","Graph CCSB")

table<-data.frame(Graph=details,Alpha=c(fitUnion$alpha,fitCCSB$alpha),"X-min"=c(fitUnion$xmin,fitCCSB$xmin),"P-Value"=c(fitUnion$KS.p,fitCCSB$KS.p))
table %>% knitr::kable(caption = "Power Law Fit Details")

```

From table 5, it can be seen that Y2H-Union and CCSB-YI network has P-Value greater than 0.05 which means that we are not able to reject the null hypothesis that the original data could have been drawn from the fitted power-law distribution. The hubs in the graphs are also significant as the $\alpha$ is between 2 and 3 in both the cases.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
##Power law fit
fitUnion <- power.law.fit(degree(lgraphUnion),9,implementation="plfit") 
fitCCSB <- power.law.fit(degree(lgraphCCSB),9,implementation="plfit") 

table<-data.frame(Graph=details,Alpha=c(fitUnion$alpha,fitCCSB$alpha),"X-min"=c(fitUnion$xmin,fitCCSB$xmin),"P-Value"=c(fitUnion$KS.p,fitCCSB$KS.p))
table %>% knitr::kable(caption = "Power Law Fit Details of largest sub-network")
```

Also from table 6, it can be seen that the largest subnetwork of Y2H-Union and CCSB-YI networks has P-Value greater than 0.05 which means that we are not able to reject the null hypothesis that the original data could have been drawn from the fitted power-law distribution. The hubs in the graphs are also significant as the $\alpha$ is between 2 and 3 in both the cases.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=50)}
data<-data.frame(prob=degree.distribution(graphUnion), degree=c(1:length(degree.distribution(graphUnion))))
data %>% ggplot(aes(x=degree,y=prob)) + geom_point() + geom_line() + labs(x="Degree",y="Probability (Degree)", title="Degree Distribution of Y2H Union PPI") +theme(plot.title = element_text(hjust = 0.5))

data<-data.frame(prob=degree.distribution(graphCCSB), degree=c(1:length(degree.distribution(graphCCSB))))
data %>% ggplot(aes(x=degree,y=prob)) + geom_point() + geom_line() + labs(x="Degree",y="Probability (Degree)", title="Degree Distribution of CCSB PPI") +theme(plot.title = element_text(hjust = 0.5))

data<-data.frame(prob=degree.distribution(lgraphUnion), degree=c(1:length(degree.distribution(lgraphUnion))))
data %>% ggplot(aes(x=degree,y=prob)) + geom_point() + geom_line() + labs(x="Degree",y="Probability (Degree)", title="Degree Distribution of Y2H Union PPI largest sub-network") +theme(plot.title = element_text(hjust = 0.5))

data<-data.frame(prob=degree.distribution(lgraphCCSB), degree=c(1:length(degree.distribution(lgraphCCSB))))
data %>% ggplot(aes(x=degree,y=prob)) + geom_point() + geom_line() + labs(x="Degree",y="Probability (Degree)", title="Degree Distribution of CCSB PPI of largest sub-network") +theme(plot.title = element_text(hjust = 0.5))
```

From the degree distribution of the plot we defines hubs which have degree greater than 10.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
hubNodesUnion<-names(which(igraph::degree(graphUnion) >= 10, arr.ind = T))
hubNodesCCSB<-names(which(igraph::degree(graphCCSB) >= 10, arr.ind = T))

rhubNodesUnion<-sample(V(graphUnion),length(hubNodesUnion))
rhubNodesCCSB<-sample(V(graphCCSB),length(hubNodesCCSB))

hubNodeslUnion<-names(which(igraph::degree(lgraphUnion) >= 10, arr.ind = T))
hubNodeslCCSB<-names(which(igraph::degree(lgraphCCSB) >= 10, arr.ind = T))

rhubNodeslUnion<-sample(V(lgraphUnion),length(hubNodeslUnion))
rhubNodeslCCSB<-sample(V(lgraphCCSB),length(hubNodeslCCSB))


#intersect(hubNodesUnion,as.vector(essentialGenes[,1]))
#Compare with the random nodes.
```


#Question 2:

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

top5percentUnion<-vcount(lgraphUnion)*0.05
centralGenesUnion<-names(sort(igraph::betweenness(lgraphUnion),decreasing = TRUE)[1:as.integer(top5percentUnion)])

top5percentCCSB<-vcount(lgraphCCSB)*0.05
centralGenesCCSB<-names(sort(igraph::betweenness(lgraphCCSB),decreasing = TRUE)[1:as.integer(top5percentCCSB)])

#intersect(hubNodesUnion,as.vector(essentialGenes[,1]))

```

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

gUniond10 <- delete.edges(lgraphUnion, E(lgraphUnion)[sample(seq_along(E(lgraphUnion)), 0.1*ecount(lgraphUnion))])

gUniond25 <- delete.edges(lgraphUnion, E(lgraphUnion)[sample(seq_along(E(lgraphUnion)), 0.25*ecount(lgraphUnion))])

gCCSBd10 <- delete.edges(lgraphCCSB, E(lgraphCCSB)[sample(seq_along(E(lgraphCCSB)), 0.1*ecount(lgraphCCSB))])

gCCSBd25 <- delete.edges(lgraphCCSB, E(lgraphCCSB)[sample(seq_along(E(lgraphCCSB)), 0.25*ecount(lgraphCCSB))])

write_graph(gUniond10,"UnionD10.txt","ncol")
write_graph(gUniond25,"UnionD25.txt","ncol")

write_graph(gCCSBd10,"CCSBD10.txt","ncol")
write_graph(gCCSBd25,"CCSBD25.txt","ncol")


```


