---
title: "BCB570 Assignment 7"
author: "Ashish Jain"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE
)
```

#5-fold cross-validation to design a classifier to separate B and T cells in ALL dataset
In this problem, first, we downloaded the ALL data from Bioconductor package. After that, we process the data and created labels for B and T cells. Then, we used the `cvpred.ksvm` function from the `kernlab` package for 5-fold cross-validation. First, we used the linear classifier for the classification.

```{r , tidy=TRUE, tidy.opts=list(width.cutoff=60)}
#source("https://bioconductor.org/biocLite.R")
#biocLite("ALL")
library(kernlab)
library(ALL)
library(ROCR)
source('general.R')
data(ALL)
# Predict type of disease from expression data
y <- substr(ALL$BT,1,1)
x <- t(exprs(ALL))

k=5
ypredscorecv <- cvpred.ksvm(x,y,folds=k,type="C-svc",kernel='vanilladot',C=1,scaled=c(),predtype="decision")

# Check the performance
print(table(ypredscorecv > 0,y))
pred <- prediction(ypredscorecv,y)
perf <- performance(pred, measure = "tpr", x.measure = "fpr")
plot(perf,main = "5-fold cross-validation for Linear classifier")
#plot(unlist(perf@x.values), unlist(perf@y.values))
```

The results show that the linear SVM classifier is able to classify both the B and T cells perfectly (with accuracy equals to 1). The AUROC is also equal to 1. Next, we also tried non-linear classifier using the `Radial Basis kernel` function. 


```{r , tidy=TRUE, tidy.opts=list(width.cutoff=60)}
##Non-linear classifier
k=5
ypredscorecv <- cvpred.ksvm(x,y,folds=k,type="C-svc",kernel='rbf',C=1,scaled=c(),predtype="decision")
# Check the performance
print(table(ypredscorecv > 0,y))
pred <- prediction(ypredscorecv,y)
perf <- performance(pred, measure = "tpr", x.measure = "fpr")
plot(perf, main = "5-fold cross-validation for Non-linear classifier")
```

The results show that the non-linear SVM classifier also detects both the B and T cells perfectly (with accuracy equals to 1). The AUROC is also equal to 1.

